---
title: "Work Sample "
author: J. Antonio García
date: "09/11/2020"
output:  pdf_document
---

```{r setup0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```
# Reduce maintenance cost through predictive techniques 

## Preamble

```{r setup, eval=FALSE}
# html_notebook

packs <- c('lubridate', 'dplyr', 'ggplot2', 'caret', 'MLmetrics', 'gbm', 'e1071', 'LiblineaR', 
           'xgboost', 'randomForest', 'doParallel')
index <- packs %in% row.names(installed.packages())
if (any(!index)){
  sapply(packs[!index], FUN=install.packages )
}
```

```{r a}
require(lubridate) # easy handle datetimes
require(dplyr) # like SQL in R, and also load pipe operator 
require(ggplot2) # easy, fast  and nice plots 
require(caret) # a toolbox 
require(MLmetrics) # metric in one line 
require(gbm) # boosting alg.
require(e1071) # numerical rutines for svm implementation 
require(LiblineaR) # another implementation for RLogReg
require(xgboost) # The implementation 
require(randomForest) # nice RF implementation 
require(doParallel) #ugly parallel in R but useful  

```


#### Set up __parallel__ enviroment

```{r parallel }
cl <- makePSOCKcluster(2)
registerDoParallel(cl)
```

# EDA

```{r eda }
rm(list=ls())  # clean env.
t1 <- Sys.time()
data.raw <- read.csv(file='device_failure.csv') # few records kernels function works fine 
print(sum(is.na(data.raw))) # NOT NULLS! THANKS A LOT :D
data.raw %>% arrange(device, date) %>% mutate(date = ymd(date) ) -> data.raw
```

Each device has 0 or 1 failure, and if has a failure it's the last row

```{r knn.fails}
n.fails.index <- which(data.raw$failure==1) #only 106 failures 
nn.fails <- data.raw[ rep(n.fails.index, each=9) + -4:4, ]
head(nn.fails, 50 )
```

And in general, the devices present at most one fault and from which no information is recorded about them. 
Above all we are in a case where the variable to predict has a __strong positive bias__, more than 99% of the records are not failures, a very common case in practice ...

```{r filter}
data.raw %>% filter(failure==1) %>% group_by(device) %>% summarise(n=n()) -> t
summary(t$n)
table(data.raw$failure) / dim(data.raw)[1]
summary(data.raw)
```

# Focus on success stories

So we decided to focus on success stories to infer from them insights that allow us to carry out the task.

```{r plots}
device.with.failures <- unique(data.raw$device[n.fails.index] )
data.sample <- data.raw[ data.raw$device %in% device.with.failures, ]
ggplot(data.sample, aes(attribute1, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal() + theme(legend.position="none") 
ggplot(data.sample, aes(attribute2, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal() +theme(legend.position="none") 
ggplot(data.sample, aes(attribute3, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal() +theme(legend.position="none") 
ggplot(data.sample, aes(attribute4, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute5, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute6, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute7, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute8, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute9, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 

```

Due to the distribution of some of the variables, we apply a non-linear transformation that allows us to more easily discriminate between failures and non-failures.


```{r aaaw}
index.columns <- c(2, 3, 4, 7, 8, 9) + 3 
# log features selected 
data.sample[, names(data.sample)[index.columns]] <-  
  log(data.sample[, names(data.sample)[index.columns]] + 1 )
# standar features 
index.columns <- grep('attr', names(data.sample))
summary(data.sample)
for ( i in index.columns ){
  temp <- data.sample[, names(data.sample)[i]]
  data.sample[, names(data.sample)[i]] <- scale(temp)
}
ggplot(data.sample, aes(attribute1, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal() + theme(legend.position="none") 
ggplot(data.sample, aes(attribute2, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal() +theme(legend.position="none") 
ggplot(data.sample, aes(attribute3, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal() +theme(legend.position="none") 
ggplot(data.sample, aes(attribute4, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute5, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute6, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute7, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute8, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 
ggplot(data.sample, aes(attribute9, fill =  as.character(failure), alpha=.01)) +
  geom_density()  + theme_minimal()+theme(legend.position="none") 

```

After we divided our sample to continue with a preselection of models, among the enormous variety of algorithms and implementations that exist, we decided to report 4, because they are interpretable models and easy to explain to non-specialized people.

Since we are interested in keeping the number of false negatives and false positives low, we opted for the $F_1$ metric to measure the performance of the algorithms.

```{r train_pre}
createPartition <- function(data_, p=0.7){
  # Inputs: data_ (data.frame) to split 
  #         p (numeric): dataframe's proportion for train sample 
  t <- unique(data_$device)
  n <- length(t)
  n.p <- round(n*p, 0)
  t.sample <- sample(t, n.p)
  train.index <- which( data_$device %in% t.sample)
  return(train.index)
}

f1 <- function (data, lev = NULL, model = NULL) {
  # Function requiere to calculate F1 score within caret::train , see doc. 
  precision <- posPredValue(data$pred, data$obs, positive = "Failure")
  recall  <- sensitivity(data$pred, data$obs, positive = "Failure")
  f1_val <- (2 * precision * recall) / (precision + recall)
  names(f1_val) <- c("F1")
  return(f1_val)
} 

set.seed(0)
data.sample$failure <- factor(data.sample$failure)
levels(data.sample$failure) <- c('NoFailure', 'Failure')
train.index <- createPartition(data.sample)
data.sample$date <- data.sample$device <- NULL
train <- data.sample[train.index, ]
test <- data.sample[-train.index, ]
fit.control <- trainControl(  method = 'repeatedcv', number = 10, repeats = 3,
                             allowParallel = TRUE, classProbs = TRUE,
                             summaryFunction = f1,  sampling =  "down")

set.seed(0)
gbmFit1 <- train(failure ~ ., data = train, method = "gbm", trControl = fit.control,
                 verbose = FALSE)
xgb.Fit1 <- train(failure ~ ., data = train, method = "xgbTree", #tuneLength = 5, search= 'random', 
                  trControl = fit.control,
                  verbose = FALSE)
rf.Fit1 <- train(failure ~ ., data = train, method = "rf", trControl = fit.control,
                 verbose = FALSE)
rlg.Fit1 <- train(failure ~ ., data = train, method = "regLogistic", 
                  trControl = fit.control, verbose = FALSE)

```

The test based on the Bonferroni intervals strongly suggests that XGB and RF perform better than the other methods, however, when evaluating it on the data test , we opted to only report RF's tunning results: 


```{r bonferroni}
resamps <- resamples(list(GBM = gbmFit1, XGB = xgb.Fit1,
                          RF = rf.Fit1, RLG=rlg.Fit1  ))
summary(resamps)
summary(diff(resamps))
t2 <- Sys.time()
t2 - t1
confusionMatrix(predict(rf.Fit1$finalModel,test), test$failure)
confusionMatrix(predict(xgb.Fit1,test), test$failure)
```

```{r tunning1}
t3 <- Sys.time()
set.seed(0)
tune_grid <- expand.grid(nrounds=c(100,300), max_depth = c(4:7), eta = c(0.05, 1),   gamma = c(0.01),
                         colsample_bytree = c(0.75), subsample = c(0.50),  min_child_weight = c(0))

xgb_fit <- train(failure ~., data = train, method = "xgbTree",
                trControl= fit.control,
                tuneGrid = tune_grid,
                tuneLength = 10)
tune_grid <- expand.grid(.mtry = (1:16)) 
rf_fit <- train(failure ~., data = train, method = "rf",
                trControl= fit.control,
                tuneGrid = tune_grid,
                tuneLength = 10)

t4 <- Sys.time()
t4 - t1
confusionMatrix(predict(rf_fit$finalModel, test), test$failure)
confusionMatrix(predict(xgb_fit, test), test$failure)

```